%%
%% This is file `thesis-ex.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% uiucthesis2009.dtx  (with options: `example')
%% 
\def\fileversion{v2.25a} \def\filedate{2009/10/10}
%% Package and Class "uiucthesis2009" for use with LaTeX2e.
\documentclass[edeposit,fullpage]{uiucthesis2009}
\usepackage[usenames, dvipsnames]{color}
\usepackage{graphicx}
\usepackage{xspace}     %to get spaces in commands right
\usepackage{subfig}
\usepackage{url}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{lipsum}
\usepackage{caption}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{amssymb}
\usepackage{cancel}
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{notoccite}
\usepackage{units}      %to get proper nobreak spaces for units
\usepackage{hyperref}   %clickable links in pdfs
\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
	language=SQL,
	aboveskip=1mm,
	belowskip=1mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numbers=none,
	numberstyle=\tiny\color{gray},
	keywordstyle=\color{blue},
	commentstyle=\color{dkgreen},
	stringstyle=\color{mauve},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3
}

\nocopyrightpage

\newcolumntype{L}{>{\hspace*{-\tabcolsep}}c}
\newcolumntype{R}{c<{\hspace*{-\tabcolsep}}}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\colorlet{tableheadcolor}{gray!25} % Table header colour = 25% gray
\newcommand{\headcol}{\rowcolor{tableheadcolor}} %
\colorlet{tablerowcolor}{gray!10} % Table row separator colour = 10% gray
\newcommand{\rowcol}{\rowcolor{tablerowcolor}}

\newcommand{\CN}{\textcolor{red}{$^{[citation\ needed]}$}}
\newcommand{\Jpsi}{J/\Psi}
\newcommand{\us}{$\mu$s\xspace}
\newcommand{\red}[1]{\emph{\textcolor{red}{#1}}}
\newcommand{\bigoh}{\mathcal{O}}
\newcommand{\dbar}{\bar{d}}
\newcommand{\ubar}{\bar{u}}
\newcommand{\sbar}{\bar{s}}
\newcommand{\cbar}{\bar{c}}

\begin{document}

\title{Nuclear dependence of proton-induced Drell-Yan dimuon production at \unit[120]{GeV} at SeaQuest}
\author{Bryan P. Dannowitz}
\department{Nuclear Physics}
\schools{B.S., New Mexico Institute of Mining and Technology, 2008}
\phdthesis
\advisor{Naomi C. R. Makins}
\degreeyear{2016}
\committee{Professor Jen-Chieh Peng, Chair\\
	Professor Naomi C. R. Makins, Director of Research\\
	Professor Michael Stone\\
	Professor Jim Eckstein}
\maketitle

\frontmatter

%% Create an abstract that can also be used for the ProQuest abstract.
%% Note that ProQuest truncates their abstracts at 350 words.
\begin{abstract}
A measurement of the atomic mass ($A$) dependence of $p+A\rightarrow\mu^+\mu^- + X$ Drell-Yan dimuons produced by \unit[120]{GeV} protons is presented here. The data was taken by the SeaQuest experiment at Fermilab using a proton beam extracted from its Main Injector. Over 61,000 dimuon pairs were recorded with invariant mass $4.2 < M_{\gamma^*} < \unit[10] GeV$ and target parton momentum fraction $0.1\leq x_2 \leq 0.5$ for nuclear targets $^1H,\ ^2H,\ C,\ Fe,\ $and$\ W$. The ratio of dimuon yields per nucleon ($Y$) for heavy nuclei versus $^2H$, $R^{DY} = Y^{(A)} / Y^{(^2H)} \approx \bar{q}_u^{(A)}(x)/\bar{q}_u^{(^2H)}(x)$, is sensitive to modifications in the anti-quark sea distributions ($\bar{q}(x)$) in nuclei for the case of proton-induced Drell-Yan. The data analyzed here provides tighter constraints on various models that attempt to define the anomalous behavior of nuclear modification as seen in deep inelastic lepton scattering, a phenomenon generally known as the EMC effect.

\end{abstract}

%% Create a dedication in italics with no heading, centered vertically
%% on the page.

\begin{dedication}
Dedicated to my grandfather, Ted, who took me to school every day.
\end{dedication}

%% Create an Acknowledgements page, many departments require you to
%% include funding support in this.
\chapter*{Acknowledgments}

This dissertation comes as a grand sum of the help and support from the many colleagues, friends, and family in my life. First and foremost of them all, I give my thanks to my parents, Judy and Robert, who have never wavered in encouraging and supporting me in every way possible as I bounded through this long journey across many states, schools, jobs, and countries.

My sincere gratitude goes to my advisor and mentor, Naomi Makins, who first drew me to the University of Illinois and SeaQuest by her enthusiasm and fervor alone. Anything insightful that I may present or state here in this paper has probably come from what I have learned through her patient and exceptionally thoughtful teachings.

Thanks also go out to my fellow UIUC graduate students have been the bedrock of my academic life here. Without the constant cameraderie of my friend Evan McClellan, I fear I would not have made it through these past years with my sanity intact, and without the tenacious, creative, problem-solving mind of Bryan Kerns, I'm nearly certain that we would not have solved as many problems at SeaQuest as we have.

Particular thanks must extended to one, Markus Diefenthaler, who by sheer force of will held SeaQuest together and kept the collaborative effort focused for so many years. Markus also played a pivotal position as a role model to Evan, Bryan, and myself as what an experimental physicist should strive to be.

I would also like to state my appreciation to the SeaQuest collaboration for placing their trust (and data!) in the hands of a humble database administrator and second-year graduate student. I have learned very much in the company of the great scientific tapestry that is the SeaQuest collaboration, and I hope to carry all of the lessons I've learned with me throughout my life. Of course, the data presented here would simply not be possible without the grand coordinated efforts by all of the collaboration members to assemble and execute such an ambitious and challenging experiment.

My thanks also go out to the staff and faculty of the Nuclear and High Energy Physics Laboratory for their availability and willingness to help a graduate student in need. In particular, the weekly feedback and insight from Jen-Chieh Peng has been simply invaluable. The assistance from Steve Errede and Allison Sibert is the only reason that the PMT testing and development got off of the ground. Finally, life at the NPL would simply grind to a halt if it were not for the constant availability, willingness, and helpfulness of Mike Suchor.

I would also like to take a moment to thank the Department of Physice, the Nuclear Physics Laboratory, and the National Science Foundation\footnote{NSF Grant No. 1506416} for providing financial support and employ as a research assistant throughout the years, thereby giving me the opportunity to do something I love for a living.

Finally, I offer my most heartfelt gratitude to my wife, Carly, for the incredible leap of faith she took moving from Hawai`i on over to the middle of the corn fields just to be with me as I put myself through the rigors of these graduate studies. Her patience, understanding, companionship, and encouragement has truly been what has kept me going all these years.

%% The thesis format requires the Table of Contents to come
%% before any other major sections, all of these sections after
%% the Table of Contents must be listed therein (i.e., use \chapter,
%% not \chapter*).  Common sections to have between the Table of
%% Contents and the main text are:
%%
%% List of Tables
%% List of Figures
%% List Symbols and/or Abbreviations
%% etc.

\tableofcontents
\listoftables
\listoffigures

%% Create a List of Abbreviations. The left column
%% is 1 inch wide and left-justified
\chapter{List of Abbreviations}

\begin{symbollist*}
	\item[ACNET] Accelerator control system network
	\item[AD] Accelerator division (of Fermilab)
	\item[BIM] Beam intensity monitor
	\item[BNC] Bayonet Neill–Concelman signal cable connector standard
	\item[BOS] Beginning of spill
	\item[BPM] Beam profile monitor
	\item[CEBAF] Continuous Electron Beam Accelerator Facility
	\item[CODA] CEBAF On-line Data Acquisition
	\item[CS] Collins-Soper reference frame
	\item[CTEQ] Coordinated theoretical-experimental project on QCD
	\item[DAQ] Data acquisition
	\item[DC] Drift chamber
	\item[DIS] Deep-inelastic scattering
	\item[DY] Drell-Yan
	\item[EOS] End of spill
	\item[EOF] End of file
	\item[EPICS] Experimental physics and industrial control system
	\item[EVIO] CODA event input/output
	\item[FEE] Front-end electronics
	\item[FNAL] Fermi National Accelerator Laboratory (Fermilab)
	\item[FPGA] Field programmable gate arrays
	\item[IC] Ion chamber
	\item[LED] Light-emitting diode
	\item[LINAC] Fermilab Linear Accelerator
	\item[LO] Leading order
	\item[MI] (Fermilab) Main Injector
	\item[MOSFET] Metal–oxide–semiconductor field-effect transistor
	\item[NDF] Neutral density filter
	\item[NLO] Next-to-leading order
	\item[NM] Neutrino-muon beam line
	\item[NNLO] Next-to-next-to-leading order
	\item[PCB] Printed circuit board
	\item[PDF] Parton distribution function
	\item[PID] Particle identification
	\item[PMT] Photomultiplier tube
	\item[RF] Radio frequency
	\item[RFQ] Radio frequency quadrupole
	\item[ROC] Readout controller
	\item[SEM] Secondary emission monitor
	\item[SHV] Secure high voltage connector standard
	\item[SLAC] Stanford Linear Accelerator Center
	\item[SRC] Short range correlations
	\item[TS] Trigger supervisor
	\item[QCD] Quantum chromodynamics
	\item[QED] Quantum electrodynamics
	\item[QIE] Charge (Q) integrator and encoder
	\item[RDBMS] Relational database management system
	\item[SQL] Structured querying language
	\item[SWIC] Segmented wire ion chamber
	\item[VME] Versa Module Europa
\end{symbollist*}

\mainmatter

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

Matters are rarely as simple as we hope them to be, and atomic nuclei are no exception. When physicists began to get a handle on proton structure, it was hoped that nuclei could be simply understood as a simple collection of the same familiar protons and neutrons. Unfortunately, it turned out that \emph{more is different}, and higher-order emergent behavior reared its head. It appeared that the distribution of partons within the proton became \emph{altered} when the proton was confined within a nucleus -- a phenomenon which is generally referred to as the EMC Effect. Since the discovery of this behavior, data and theories have proceeded to accumulate for over 30 years without a completely satisfying explanation, though recent breakthroughs give a reason for optimism.

The SeaQuest experiment was commissioned and began taking data in 2012 with the expressed purpose to, among other physics goals, shed some additional light on this phenomenon. I joined the collaboration in 2009 and have had the distinct pleasure of being a part of a large-scale, medium-energy experiment as it went from being an empty hall to a testing ground for nuclear physics. It has truly been a rewarding experience being able to contribute to a coordinated, monumental effort from the point of assembly all the way to its first publishable results (or at least very close). In this thesis, I will describe in detail these contributions to the experiment. 

Chapter 1 will provide a broad outline regarding the background to the underlying physics and phenomenology. I describe the state of the field regarding the physics of unpolarized sea quark distributions along with what SeaQuest can contribute. Emphasis is placed on the phenomenon of modification of the parton distribution functions when in the presence of a nuclear medium.

In Chapter 2, the apparatus of the SeaQuest experiment is laid out, briefly covering every significant aspect, including the beam, targets, magnets, detectors, and data acquisition systems. For further information on my contributions to the SeaQuest experiment's apparatus and operations, I have laid out in the following chapters my work on the curation of the collaboration's data productions (Chapter 3) and my work upgrading the hodoscope photomultiplier tube bases for improved performance (Chapter 4). These account for a significant amount of my own time and effort on the experiment but are \emph{not} required for understanding the background and analysis regarding the primary measurement of this paper.

Chapter 5 summarizes how the data used in these studies are reconstructed and selected. This includes a brief overview of the tracking algorithm that reconstructs tracks and dimuons from the raw detector readouts, and then a thorough rundown of the data reduction, showing in great detail how good events are chosen from the whole set.

In Chapter 6, I step through the many stages of the analysis, going from the raw reconstructed data all the way through the target-to-target normalization and a litany of corrections in order to arrive at a final result. Particular emphasis is placed on the corrections pertaining to the rate dependence, which has been the key obstacle in preparing preliminary results for all SeaQuest analyses. This section goes perhaps into more detail than needed, which is intentional, as this document is, in part, meant to act as an explanatory road map for future students and researchers who wish to perform this same analysis on the full SeaQuest dataset in the future.

Chapter 7 proceeds to summarize the fully processed and corrected findings, providing a comprehensive review of $R^{DY}$, the per-nucleon Drell-Yan cross section ratio of heavy nuclei to deuterium, against various kinematics. Results are briefly discussed and compared against theoretical predictions, though more refined corrections and more statistics from further SeaQuest data taking are required to come to any firm conclusions.

\include{background}
\include{methods}
\include{productions}
\include{pmtupgrade}
\include{tracking}
\include{analysis}
\include{results}
%\include{discussion}
\include{conclusions}

\appendix

%\include{Appendix}

\backmatter

\bibliographystyle{ieeetr}
\bibliography{thesisbib}


\end{document}
\endinput
%%
